3D Magnetic modelling of ellipsoidal bodies

### 2017-02-23 manuscript/manuscript.tex
conclusion

Another particularity of ellipsoids is that they are
the only bodies which enable an analytical computation 
of its self-demagnetization.
The self-demagnetization contributes to decrease the
magnitude of the magnetization along the shortest 
axes of a body.
It is a function of the body shape and gives rise
to shape anisotropy \citep{uyeda1963, thompson1986,
dunlop1997, clark1999, tauxe2003}.
It is well-established in the literature that
the self-demagnetization can
be neglected if the body has a susceptibility lower 
than 0.1 SI 
\citep{emerson1985, clark1986, eskola1980, guo1998,
guo2001, purss2005, hillan2013, austin2014, clark2014}.
However, neglecting the self-demagnetization in
geological bodies with high susceptibilities (> 0.1 SI) 
may strongly mislead the interpretation obtained from
magnetic methods.
\citet{farrar1979} demonstrated the importance of the
ellipsoidal model in taking into account the 
self-demagnetization and determining reliable 
drilling directions on the Tennant Creek field, 
Australia.
Posteriorly, \citet{hoschke1991} also showed how the
ellipsoidal model proved to be highly successful in
locating and defining ironstone bodies in the
Tennant Creek field.
\citet{clark2000} provides a good discussion about the
influence of the self-demagnetization in magnetic
interpretation of the Osborne copper-gold deposit, 
Australia. This deposit is hosted by ironstone 
bodies that have very high susceptibility.
According to \citet{clark2000}, neglecting the effects 
of self-demagnetisation led errors of $\approx 55^{\circ}$
in the interpreted dip. 
Recently, \citet{austin2014} used magnetic modelling 
and rock property measurements to show that,
contrary to previous interpretations, the magnetization
of the Candelaria iron oxide copper-gold deposit, 
Chile, is not dominated by the induced component.
Rather, the deposit has a relatively weak remanent 
magnetization and is strongly affected by 
self-demagnetization.
These examples show the importance of the self-demagnetization 
and the ellipsoidal model in 
producing trustworthy geological models of
high-susceptibility orebodies, which may save significant 
cost associated with drilling.

A vast literature about the magnetic
modelling of ellipsoidal bodies was developed in which 
are to be found the names of many brilliant researchers.
Nevertheless, interest in this subject has not yet died
out, as is evidenced by a list of modern papers in this
field.
Besides, the geoscientific community lacks an easy-to-use 
tool to simulate the magnetic field produced by uniformly 
magnetized ellipsoids.
Such a tool could prove to be useful either for teaching
and researching geophysics.
To fill this gap, we provide here a set of routines to 
model the magnetic field produced by ellipsoids.
The routines are written in Python language as part of
the Fatiando a Terra \citep{uieda2013}, which is an 
open-source library for modelling and inversion in geophysics.
We attempt to use the best practices of continuous 
integration, documentation, unit-testing, and version-control
for the purpose of providing a reliable and easy-to-use
code.
It is hoped that the presentation of the theoretical and
practical aspects given here will be found useful to the
wide community of geoscientists.

@@

Let $\tilde{\mathbf{N}}^{\dagger}$ be the transformed
depolarization tensor calculated for the case in which
$\tilde{\mathbf{r}}$ (Eq. \ref{eq:coord_transformation})
represents a point located outside
the ellipsoidal body. In this case, 
the elements of $\tilde{\mathbf{N}}^{\dagger}$
are calculated according to Eq. \ref{eq:nij-tilde}, 
with $\tilde{f}(\tilde{\mathbf{r}})$ given by 
$\tilde{f}^{\dagger}(\tilde{\mathbf{r}})$ (Eq. \ref{eq:fi-tilde}).
As we have already pointed out, the  
$\tilde{f}^{\dagger}(\tilde{\mathbf{r}})$ (Eq. \ref{eq:fi-tilde}) is a
quadratic function of the spatial coordinates $\tilde{x}$,
$\tilde{y}$ and $\tilde{z}$. Consequently, the 
elements $\tilde{n}^{\dagger}_{ij}$, $i = 1, 2, 3$, 
$j = 1, 2, 3$, of $\tilde{\mathbf{N}}^{\dagger}$ 
do not depend on the elements of the transformed position
vector $\tilde{\mathbf{r}}$ (Eq. \ref{eq:coord_transformation}). 
Besides, the off-diagonal elements are zero and
the diagonal elements are given by \citep{stoner1945}:
\begin{equation}
\tilde{n}^{\dagger}_{ii} = \frac{abc}{2}
\int_{0}^{\infty} \frac{1}{\left( e_{i}^{2} 
+ u \right) R(u)} \, du \: , \quad i = 1, 2, 3 \: ,
\label{eq:n-tilde-dagger-ii}
\end{equation}
where $R(u)$ is defined by Eq. \ref{eq:R} and
$e_{1} = a$, $e_{2} = b$, and $e_{3} = c$. These elements 
are commonly known as \textit{demagnetizing factors}
and are defined according to the ellipsoid type.
Here, we calculate the demagnetizing factors in the SI
system. Consequently, they satisfies the condition
$\tilde{n}^{\dagger}_{11} + \tilde{n}^{\dagger}_{22} + 
\tilde{n}^{\dagger}_{33} = 1$, independently of the
ellipsoid type.
It is worth stressing that, according to Eq. \ref{eq:n-tilde-dagger-ii},
the demagnetizing factors $\tilde{n}^{\dagger}_{ii}$ are 
constants defined by the ellipsoid semi-axes $a$, $b$, and $c$.

@@

%The angular difference $\theta$ between the resultant magnetization
%and the inducing field is the same, whether the quantities
%are defined in the main coordinate system (Fig. \ref{fig:orientation-angles}a)
%or local coordinate system (Fig. \ref{fig:orientation-angles}b).
%For convenience, we define $\theta$ (in degrees) in the main coordinate system
%by using the scalar product definition as follows:
%\begin{equation}
%\theta = \frac{180}{\pi} \, \cos^{-1} \left( 
%\frac{\mathbf{M}^{\top}\mathbf{H}_{0}}
%{\| \mathbf{M} \| \; \| \mathbf{H}_{0} \|}\right) \: ,
%\label{eq:theta}
%\end{equation}
%where $\mathbf{M}$ is obtained by pre-multiplying the
%magnetization $\tilde{\mathbf{M}}$ (Eq. \ref{eq:M-tilde-remanence})
%by the orthogonal matrix $\mathbf{V}$ (Eq. \ref{eq:A}),
%which is defined according to the ellipsoid type, and
%$\mathbf{H}_{0}$ is given by Eq. \ref{eq:H0}.
%
For the particular case in which the susceptibility is isotropic,
the susceptibility tensor is defined according to Eq. \ref{eq:K-isotropic}.
In this case, the magnetization $\mathbf{M}$ (Eqs. \ref{eq:M-KH} and \ref{eq:M-remanence}),
referred to the main coordinate system (Fig. \ref{fig:orientation-angles}a),
and the matrix $\mathbf{\Lambda}$ (Eq. \ref{eq:Lambda})
can be rewritten as follows:
\begin{equation}
\mathbf{M} = \mathbf{\Lambda} \left( \chi \, \mathbf{H}_{0} +
\mathbf{M}_{R} \right) \: ,
\label{eq:M-K-isotropic}
\end{equation}
and
\begin{equation}
\mathbf{\Lambda} = \mathbf{V}
\left( \mathbf{I} + \chi \, \tilde{\mathbf{N}}^{\dagger} \right)^{-1}
\mathbf{V}^{\top} \: .
\label{eq:Lambda-K-isotropic}
\end{equation}
Despite the coordinate transformation represented by
matrix $\mathbf{V}$ (Eq. \ref{eq:A}),
this equation is in perfect agreement with those presented 
by \citet[Eqs. ~13--15]{guo2001}.
The first term, depending on the inducing field $\mathbf{H}_{0}$
(Eq. \ref{eq:H0}), 
represents the induced magnetization whereas the term depending on
$\mathbf{M}_{R}$ is the remanent magnetization.
Equation \ref{eq:M-K-isotropic} reveals that, as pointed out by
many authors
\citep[e.g.,][]{maxwell1873, dubois1896, stoner1945, clark1986, sttraton2007},
the induced magnetization opposes the inducing field 
if it is parallel to an ellipsoid axis, 
independently of the ellipsoid type. Otherwise, the
magnetization is not necessarily parallel to
the inducing field.
If we additionally consider that $\chi << 1$, the matrix 
$\mathbf{\Lambda}$ (Eq. \ref{eq:Lambda-K-isotropic})
approaches to the identity and the magnetization
$\mathbf{M}$ (Eq. \ref{eq:M-K-isotropic})
can be approximated by:
\begin{equation}
\breve{\mathbf{M}} = \chi \, \mathbf{H}_{0} +
\mathbf{M}_{R} \: ,
\label{eq:M-approx}
\end{equation}
which is the classical equation describing the resultant
magnetization in applied geophysics \citep[p. ~89]{blakely1996}.
Notice that, in this particular case, the induced magnetization
is parallel to the inducing field $\mathbf{H}_{0}$ (Eq. \ref{eq:H0}),
whether it is parallel to an ellipsoid axis or not.
Usually, Eq. \ref{eq:M-approx}
is considered a good approximation for $\chi \leq 0.1$ SI.
Although this value has been largely used in the literature,
there have been few empirical and/or theoretical investigations 
about it.

@@

By using the concept of vector norm and its corresponding
operator norm \citep{demmel1997, golub2013}, we may use Eq.
\ref{eq:perturbed-M-system} to write the following inequality:
\begin{equation}
\frac{\| \delta \mathbf{M} \|}{\| \mathbf{M} \|} \leq
\| \delta \mathbf{\Lambda}^{-1} \| \: .
\label{eq:perturbed-M-inequality}
\end{equation}
where $\| \delta \mathbf{M} \|$ and $\| \mathbf{M} \|$
denote Euclidean norms (or 2-norms) and the term 
$\| \delta \mathbf{\Lambda}^{-1} \|$ denotes the matrix 2-norm
of $\delta \mathbf{\Lambda}^{-1}$.
By using Eqs. \ref{eq:inverse-Lambda} and \ref{eq:perturbed-inverse-Lambda}
and the orthogonal invariance of the matrix 2-norm
\citep{demmel1997, golub2013}, we define
$\| \delta \mathbf{\Lambda}^{-1} \|$ as follows:
\begin{equation}
\| \delta \mathbf{\Lambda}^{-1} \| = \chi \,
\tilde{n}^{\dagger}_{max} \: ,
\label{eq:norm-perturbed-inverse-Lambda}
\end{equation}
where $\tilde{n}^{\dagger}_{max}$ is the demagnetization
factor associated to the shortest ellipsoid semi-axis.
For a triaxial ellipsoid, 
$\tilde{n}^{\dagger}_{max} \equiv \tilde{n}^{\dagger}_{33}$
(Eq. \ref{eq:n-tilde-dagger-33-triaxial}). For a prolate ellipsoid,
$\tilde{n}^{\dagger}_{max} \equiv \tilde{n}^{\dagger}_{22}$
(Eq. \ref{eq:n-tilde-dagger-22-prolate}). 
For an oblate ellipsoid,
$\tilde{n}^{\dagger}_{max} \equiv \tilde{n}^{\dagger}_{11}$
(Eq. \ref{eq:n-tilde-dagger-11-oblate}).
It is worth stressing that, independently of the ellipsoid
type, $\tilde{n}^{\dagger}_{max}$ is
a scalar function of the ellipsoid semi-axes.
In Eq. \ref{eq:perturbed-M-inequality}, the ratio
$\| \delta \mathbf{M} \| \, \| \mathbf{M} \|^{-1}$ represents the
\textit{relative error} in the approximated magnetization
$\breve{\mathbf{M}}$ (Eq. \ref{eq:M-approx}) with respect
to the true magnetization $\mathbf{M}$ (Eqs. \ref{eq:M-K-isotropic}
and \ref{eq:M-K-isotropic-inverse-Lambda}).

Given a target relative error
$\epsilon = \| \delta \mathbf{M} \| \, \| \mathbf{M} \|^{-1}$
and an ellipsoid
with given semi-axes, we may use the inequality
represented by Eq. \ref{eq:norm-perturbed-inverse-Lambda}
to define
\begin{equation}
\chi_{max} = \frac{\epsilon}{\tilde{n}^{\dagger}_{max}} \: ,
\label{eq:chi-max}
\end{equation}
which represents the maximum isotropic susceptibility 
that the ellipsoidal body can assume in order o guarantee
a relative error lower than or equal to $\epsilon$.
For isotropic susceptibilities greater than $\chi_{max}$,
there is no guarantee that the relative error in the
approximated magnetization $\breve{\mathbf{M}}$ (Eq. \ref{eq:M-approx})
with respect to the true magnetization $\mathbf{M}$ 
(Eqs. \ref{eq:M-K-isotropic} and \ref{eq:M-K-isotropic-inverse-Lambda})
is lower than or equal to $\epsilon$. The geoscientific community
has been using $\chi_{max} = 0.1$ SI as a limit value for
neglecting the self-demagnetization and, consequently,
use magnetization $\breve{\mathbf{M}}$ (Eq. \ref{eq:M-approx})
as a good approximation of the true magnetization $\mathbf{M}$ 
(Eqs. \ref{eq:M-K-isotropic} and \ref{eq:M-K-isotropic-inverse-Lambda}).
Equation \ref{eq:chi-max}, on the other hand, defines $\chi_{max}$ 
as a function of the ellipsoid semi-axes, according to a
user-specified relative error $\epsilon$.

@@

\subsection{Relative error in neglecting the self-demagnetization}

@@

\section{Conclusions}  %% \conclusions[modified heading if necessary]
TEXT

@@

The code is freely distributed under 
BSD 3-clause licence as part of the Fatiando a Terra 
\citep{uieda2013}, which is an open-source Python 
library for modelling and inversion in geophysics.
Documentation and installation instructions for 
the most current release version of Fatiando a Terra 
are provided at \url http://www.fatiando.org.

>>>>

Another particularity of ellipsoids is that they are
the only bodies which enable an analytical computation 
of its self-demagnetization.
The self-demagnetization contributes to decrease the
magnitude of the magnetization along the shortest 
axes of a body.
It is a function of the body shape and gives rise
to shape anisotropy \citep{uyeda1963, thompson1986,
dunlop1997, clark1999, tauxe2003}.
It is well-established in the literature that
the self-demagnetization can be neglected if the 
body has a susceptibility lower than 0.1 \unit{SI} 
\citep{emerson1985, clark1986, eskola1980, guo1998,
guo2001, purss2005, hillan2013, austin2014, clark2014}.
On the other hand, neglecting the self-demagnetization in
geological bodies with high susceptibilities (> 0.1 \unit{SI}) 
may strongly mislead the interpretation obtained from
magnetic methods.
This limiting value, however, seems to be determined empirically
and, so far, there has been little discussion about
how it was determined.

\citet{farrar1979} demonstrated the importance of the
ellipsoidal model in taking into account the 
self-demagnetization and determining reliable 
drilling directions on the Tennant Creek field, 
Australia.
Posteriorly, \citet{hoschke1991} also showed how the
ellipsoidal model proved to be highly successful in
locating and defining ironstone bodies in the
Tennant Creek field.
\citet{clark2000} provides a good discussion about the
influence of the self-demagnetization in magnetic
interpretation of the Osborne copper-gold deposit, 
Australia. This deposit is hosted by ironstone 
bodies that have very high susceptibility.
According to \citet{clark2000}, neglecting the effects 
of self-demagnetisation led errors of $\approx 55^{\circ}$
in the interpreted dip. 
Recently, \citet{austin2014} used magnetic modelling 
and rock property measurements to show that,
contrary to previous interpretations, the magnetization
of the Candelaria iron oxide copper-gold deposit, 
Chile, is not dominated by the induced component.
Rather, the deposit has a relatively weak remanent 
magnetization and is strongly affected by 
self-demagnetization.
These examples show the importance of the self-demagnetization 
and the ellipsoidal model in 
producing trustworthy geological models of
high-susceptibility orebodies, which may save significant 
cost associated with drilling.

A vast literature about the magnetic
modelling of ellipsoidal bodies was developed in which 
are to be found the names of many researchers.
Nevertheless, interest in this subject has not yet died
out, as is evidenced by a list of modern papers in this
field.
Besides, the geoscientific community lacks an easy-to-use 
tool to simulate the magnetic field produced by uniformly 
magnetized ellipsoids.
Such a tool could prove to be useful either for teaching
and researching geophysics.

In this work, we present a review of the vast literature
about the magnetic modelling of ellipsoidal bodies and a
theoretical discussion about the determination of the 
isotropic susceptibility value above which
the self-demagnetization must be taken into consideration.
We propose an alternative way of determining this value
based on the body shape and the maximum relative error
allowed in the resultant magnetization.
This alternative approach is validated
by the results obtained with numerical simulations.
We also provide a set of routines to 
model the magnetic field produced by ellipsoids.
The routines are written in Python language as part of
the Fatiando a Terra \citep{uieda2013}, which is an 
open-source library for modelling and inversion in geophysics.
We attempt to use the best practices of continuous 
integration, documentation, unit-testing, and version-control
for the purpose of providing a reliable and easy-to-use
code.

@@

Let $\tilde{\mathbf{N}}^{\dagger}$ be the transformed
depolarization tensor calculated for the case in which
$\tilde{\mathbf{r}}$ (Eq. \ref{eq:coord_transformation})
represents a point located outside
the ellipsoidal body. In this case, 
the elements of $\tilde{\mathbf{N}}^{\dagger}$
are calculated according to Eq. \ref{eq:nij-tilde}, 
with $\tilde{f}(\tilde{\mathbf{r}})$ given by 
$\tilde{f}^{\dagger}(\tilde{\mathbf{r}})$ (Eq. \ref{eq:fi-tilde}).
As we have already pointed out, the  
$\tilde{f}^{\dagger}(\tilde{\mathbf{r}})$ (Eq. \ref{eq:fi-tilde}) is a
quadratic function of the spatial coordinates $\tilde{x}$,
$\tilde{y}$ and $\tilde{z}$. Consequently, the 
elements $\tilde{n}^{\dagger}_{ij}$, $i = 1, 2, 3$, 
$j = 1, 2, 3$, of $\tilde{\mathbf{N}}^{\dagger}$ 
do not depend on the elements of the transformed position
vector $\tilde{\mathbf{r}}$ (Eq. \ref{eq:coord_transformation}). 
Besides, the off-diagonal elements are zero and
the diagonal elements are given by \citep{stoner1945}:
\begin{equation}
\tilde{n}^{\dagger}_{ii} = \frac{abc}{2}
\int_{0}^{\infty} \frac{1}{\left( e_{i}^{2} 
+ u \right) R(u)} \, du \: , \quad i = 1, 2, 3 \: ,
\label{eq:n-tilde-dagger-ii}
\end{equation}
where $R(u)$ is defined by Eq. \ref{eq:R} and
$e_{1} = a$, $e_{2} = b$, and $e_{3} = c$. These elements 
are commonly known as \textit{demagnetizing factors}
and are defined according to the ellipsoid type.
Here, we calculate the demagnetizing factors in the \unit{SI}
system. Consequently, they satisfies the condition
$\tilde{n}^{\dagger}_{11} + \tilde{n}^{\dagger}_{22} + 
\tilde{n}^{\dagger}_{33} = 1$, independently of the
ellipsoid type.
It is worth stressing that, according to Eq. \ref{eq:n-tilde-dagger-ii},
the demagnetizing factors $\tilde{n}^{\dagger}_{ii}$ are 
constants defined by the ellipsoid semi-axes $a$, $b$, and $c$.

@@

%The angular difference $\theta$ between the resultant magnetization
%and the inducing field is the same, whether the quantities
%are defined in the main coordinate system (Fig. \ref{fig:orientation-angles}a)
%or local coordinate system (Fig. \ref{fig:orientation-angles}b).
%For convenience, we define $\theta$ (in degrees) in the main coordinate system
%by using the scalar product definition as follows:
%\begin{equation}
%\theta = \frac{180}{\pi} \, \cos^{-1} \left( 
%\frac{\mathbf{M}^{\top}\mathbf{H}_{0}}
%{\| \mathbf{M} \| \; \| \mathbf{H}_{0} \|}\right) \: ,
%\label{eq:theta}
%\end{equation}
%where $\mathbf{M}$ is obtained by pre-multiplying the
%magnetization $\tilde{\mathbf{M}}$ (Eq. \ref{eq:M-tilde-remanence})
%by the orthogonal matrix $\mathbf{V}$ (Eq. \ref{eq:A}),
%which is defined according to the ellipsoid type, and
%$\mathbf{H}_{0}$ is given by Eq. \ref{eq:H0}.
%
For the particular case in which the susceptibility is isotropic,
the susceptibility tensor is defined according to Eq. \ref{eq:K-isotropic}.
In this case, the magnetization $\mathbf{M}$ (Eqs. \ref{eq:M-KH} and \ref{eq:M-remanence}),
referred to the main coordinate system (Fig. \ref{fig:orientation-angles}a),
and the matrix $\mathbf{\Lambda}$ (Eq. \ref{eq:Lambda})
can be rewritten as follows:
\begin{equation}
\mathbf{M} = \mathbf{\Lambda} \left( \chi \, \mathbf{H}_{0} +
\mathbf{M}_{R} \right) \: ,
\label{eq:M-K-isotropic}
\end{equation}
and
\begin{equation}
\mathbf{\Lambda} = \mathbf{V}
\left( \mathbf{I} + \chi \, \tilde{\mathbf{N}}^{\dagger} \right)^{-1}
\mathbf{V}^{\top} \: .
\label{eq:Lambda-K-isotropic}
\end{equation}
Despite the coordinate transformation represented by
matrix $\mathbf{V}$ (Eq. \ref{eq:A}),
this equation is in perfect agreement with those presented 
by \citet[Eqs. ~13--15]{guo2001}.
The first term, depending on the inducing field $\mathbf{H}_{0}$
(Eq. \ref{eq:H0}), 
represents the induced magnetization whereas the term depending on
$\mathbf{M}_{R}$ is the remanent magnetization.
Equation \ref{eq:M-K-isotropic} reveals that, as pointed out by
many authors
\citep[e.g.,][]{maxwell1873, dubois1896, stoner1945, clark1986, sttraton2007},
the induced magnetization opposes the inducing field 
if it is parallel to an ellipsoid axis, 
independently of the ellipsoid type. Otherwise, the
magnetization is not necessarily parallel to
the inducing field.
If we additionally consider that $\chi << 1$, the matrix 
$\mathbf{\Lambda}$ (Eq. \ref{eq:Lambda-K-isotropic})
approaches to the identity and the magnetization
$\mathbf{M}$ (Eq. \ref{eq:M-K-isotropic})
can be approximated by:
\begin{equation}
\breve{\mathbf{M}} = \chi \, \mathbf{H}_{0} +
\mathbf{M}_{R} \: ,
\label{eq:M-approx}
\end{equation}
which is the classical equation describing the resultant
magnetization in applied geophysics \citep[p. ~89]{blakely1996}.
Notice that, in this particular case, the induced magnetization
is parallel to the inducing field $\mathbf{H}_{0}$ (Eq. \ref{eq:H0}),
whether it is parallel to an ellipsoid axis or not.
Usually, Eq. \ref{eq:M-approx}
is considered a good approximation for $\chi \leq 0.1$ \unit{SI}.
Although this value has been largely used in the literature,
there have been few empirical and/or theoretical investigations 
about it.

@@

By using the concept of vector norm and its corresponding
operator norm \citep{demmel1997, golub2013}, we may use Eq.
\ref{eq:perturbed-M-system} to write the following inequality:
\begin{equation}
\frac{\| \delta \mathbf{M} \|}{\| \mathbf{M} \|} \leq
\| \delta \mathbf{\Lambda}^{-1} \| \: .
\label{eq:perturbed-M-inequality}
\end{equation}
where $\| \delta \mathbf{M} \|$ and $\| \mathbf{M} \|$
denote Euclidean norms (or 2-norms) and the term 
$\| \delta \mathbf{\Lambda}^{-1} \|$ denotes the matrix 2-norm
of $\delta \mathbf{\Lambda}^{-1}$.
By using Eqs. \ref{eq:inverse-Lambda} and \ref{eq:perturbed-inverse-Lambda}
and the orthogonal invariance of the matrix 2-norm
\citep{demmel1997, golub2013}, we define
$\| \delta \mathbf{\Lambda}^{-1} \|$ as follows:
\begin{equation}
\| \delta \mathbf{\Lambda}^{-1} \| = \chi \,
\tilde{n}^{\dagger}_{max} \: ,
\label{eq:norm-perturbed-inverse-Lambda}
\end{equation}
where $\tilde{n}^{\dagger}_{max}$ is the demagnetization
factor associated to the shortest ellipsoid semi-axis.
For a triaxial ellipsoid, 
$\tilde{n}^{\dagger}_{max} \equiv \tilde{n}^{\dagger}_{33}$
(Eq. \ref{eq:n-tilde-dagger-33-triaxial}), for a prolate ellipsoid,
$\tilde{n}^{\dagger}_{max} \equiv \tilde{n}^{\dagger}_{22}$
(Eq. \ref{eq:n-tilde-dagger-22-prolate}), and,
for an oblate ellipsoid,
$\tilde{n}^{\dagger}_{max} \equiv \tilde{n}^{\dagger}_{11}$
(Eq. \ref{eq:n-tilde-dagger-11-oblate}).
It is worth stressing that, independently of the ellipsoid
type, $\tilde{n}^{\dagger}_{max}$ is
a scalar function of the ellipsoid semi-axes.
In Eq. \ref{eq:perturbed-M-inequality}, the ratio
$\| \delta \mathbf{M} \| \, \| \mathbf{M} \|^{-1}$ represents the
\textit{relative error} in the approximated magnetization
$\breve{\mathbf{M}}$ (Eq. \ref{eq:M-approx}) with respect
to the true magnetization $\mathbf{M}$ (Eqs. \ref{eq:M-K-isotropic}
and \ref{eq:M-K-isotropic-inverse-Lambda}).
Given a target relative error $\epsilon$ and an ellipsoid
with given semi-axes, we may use the inequality
represented by Eq. \ref{eq:norm-perturbed-inverse-Lambda}
to define
\begin{equation}
\chi_{max} = \frac{\epsilon}{\tilde{n}^{\dagger}_{max}} \: ,
\label{eq:chi-max}
\end{equation}
which represents the maximum isotropic susceptibility 
that the ellipsoidal body can assume in order o guarantee
a relative error lower than or equal to $\epsilon$.
For isotropic susceptibilities greater than $\chi_{max}$,
there is no guarantee that the relative error in the
approximated magnetization $\breve{\mathbf{M}}$ (Eq. \ref{eq:M-approx})
with respect to the true magnetization $\mathbf{M}$ 
(Eqs. \ref{eq:M-K-isotropic} and \ref{eq:M-K-isotropic-inverse-Lambda})
is lower than or equal to $\epsilon$. The geoscientific community
has been using $\chi_{max} = 0.1$ \unit{SI} as a limit value for
neglecting the self-demagnetization and, consequently,
use magnetization $\breve{\mathbf{M}}$ (Eq. \ref{eq:M-approx})
as a good approximation of the true magnetization $\mathbf{M}$ 
(Eqs. \ref{eq:M-K-isotropic} and \ref{eq:M-K-isotropic-inverse-Lambda}).
Equation \ref{eq:chi-max}, on the other hand, defines $\chi_{max}$ 
as a function of the ellipsoid semi-axes, according to a
user-specified relative error $\epsilon$.

@@

@@

\citet{farrar1979} represented the Warrego orebody as a triaxial
ellipsoid having a high isotropic susceptibility.
In this case, the self-demagnetization strongly impacts the
magnetic modelling of this body.
According to \citet{wedekind1990}, the Warrego orebody is a combination of
two major and several small ironstone lodes, which are discrete bodies
comprised predominantly of magnetite or hematite above the base of
oxidation.
The Warrego orebody was the resource on which the Warrego mine was
developed. After nearly a decade as one of the most important 
gold and copper mines in Australia, the Warrego mine was closed 
in late 1989.

Table \ref{tab:parameters-warrego} shows the parameters defining
a synthetic orebody which is based on that presented by
\citet{farrar1979} to represent the Warrego orebody.
Figure \ref{fig:tfa-warrego} shows the total-field anomaly
$\Delta T (\mathbf{r})$ (Eq. \ref{eq:delta-T-approx}) produced
by the synthetic body on a regular grid of $100 \times 100$
points at a constant vertical coordinate $z = 0$ \unit{m}.
The total-field anomaly varies from $\approx -295.3$ \unit{nT}
to $\approx 149.5$ \unit{nT}, resulting in a peak-to-peak amplitude
of $\approx 444.9$ \unit{nT}, and was calculated by using the 
true magnetization $\mathbf{M}$ defined in Eqs. 
\ref{eq:M-K-isotropic} and \ref{eq:M-K-isotropic-inverse-Lambda}.

We have calculated the difference between the total-field anomaly
$\Delta T (\mathbf{r})$ (Eq. \ref{eq:delta-T-approx}) calculated with the
true magnetization $\mathbf{M}$ (Eqs. 
\ref{eq:M-K-isotropic} and \ref{eq:M-K-isotropic-inverse-Lambda})
and that calculated with the approximated magnetization
$\breve{\mathbf{M}}$ (Eq. \ref{eq:M-approx}).
The differences were calculated by using the synthetic body 
defined in Tab. \ref{tab:parameters-warrego}, but with three
different isotropic susceptibilities.
Figures \ref{fig:tfa-differences-warrego}a, \ref{fig:tfa-differences-warrego}b
and \ref{fig:tfa-differences-warrego}c show the differences calculated
by using, respectively, isotropic susceptibilities $\chi = 1.69$ \unit{SI} 
(Tab. \ref{tab:parameters-warrego}), $\chi_{1} = 0.1$ \unit{SI} and 
$\chi_{2} = 0.116$ \unit{SI}.

As expected, the differences calculated by using the higher isotropic
susceptibility (Fig. \ref{fig:tfa-differences-warrego}a) are very large.
The peak-to-peak amplitude is $\approx 212.0$ \unit{nT}
and represents $\approx 48 \%$ of the peak-to-peak amplitude of
the total-field anomaly shown in Fig. \ref{fig:tfa-warrego}.
This result reinforces that the use of the approximated magnetization
$\breve{\mathbf{M}}$ (Eq. \ref{eq:M-approx}) may negatively impact
the magnetic modelling of bodies with high isotropic susceptibility.

Figure \ref{fig:tfa-differences-warrego}b shows the differences calculated
by using $\chi_{1} = 0.1$ \unit{SI}.
It is commonly accepted that, for bodies having isotropic susceptibilities
lower than or equal to $0.1$ \unit{SI}, the self-demagnetization can be neglected and,
consequently, the magnetization $\breve{\mathbf{M}}$ (Eq. \ref{eq:M-approx})
is a good approximation of the true magnetization $\mathbf{M}$
(Eqs. \ref{eq:M-K-isotropic} and \ref{eq:M-K-isotropic-inverse-Lambda}).
In our test, the use of $\chi_{1} = 0.1$ \unit{SI} leads to a relative error
$\| \delta \mathbf{M} \| \, \| \mathbf{M} \|^{-1} \approx 4.2 \%$
(Eq. \ref{eq:perturbed-M-inequality}) in the magnetization.
The peak-to-peak amplitude of the differences in the total-field anomaly
(Fig. \ref{fig:tfa-differences-warrego}b) is $\approx 1.5$ \unit{nT},
which represents $\approx 4.5 \%$ of the peak-to-peak amplitude of the
total-field anomaly calculates by using the true magnetization $\mathbf{M}$
(Eqs. \ref{eq:M-K-isotropic} and \ref{eq:M-K-isotropic-inverse-Lambda}).

Finally, Fig. \ref{fig:tfa-differences-warrego}c shows the differences
calculated by using $\chi_{2} = 0.116$ \unit{SI}.
This value was calculated by using Eq. \ref{eq:chi-max} with a
target relative error $\epsilon = 8 \%$ and the $\tilde{n}^{\dagger}_{max}$
defined by Eq. \ref{eq:n-tilde-dagger-33-triaxial}.
By using this isotropic susceptibility, it is expected that the
calculated relative error 
$\| \delta \mathbf{M} \| \, \| \mathbf{M} \|^{-1}$
(Eq. \ref{eq:perturbed-M-inequality}) in the magnetization be
lower than or equal to the target relative error $\epsilon = 8 \%$.
In this test, the use of $\chi_{2} = 0.116$ \unit{SI} leads to a relative error
$\| \delta \mathbf{M} \| \, \| \mathbf{M} \|^{-1} \approx 4.9 \%$
(Eq. \ref{eq:perturbed-M-inequality}) in the magnetization.
The peak-to-peak amplitude of the differences in the total-field anomaly
(Fig. \ref{fig:tfa-differences-warrego}c) is $\approx 2.0$ \unit{nT},
which represents $\approx 5.1 \%$ of the peak-to-peak amplitude of the
total-field anomaly calculates by using the true magnetization $\mathbf{M}$
(Eqs. \ref{eq:M-K-isotropic} and \ref{eq:M-K-isotropic-inverse-Lambda}).
In this case, the use an isotropic susceptibility greater than the usual
limit $0.1$ \unit{SI} does not mislead the magnetic modelling
dramatically. On the contrary, it shows small discrepancies in the
magnetic modelling and validates Eq. \ref{eq:chi-max}.

\section{Conclusions}  %% \conclusions[modified heading if necessary]

We present an integrated review of the vast literature 
about the magnetic modelling of triaxial, prolate and
oblate ellipsoids.
We also present a theoretical discussion about the
determination of the isotropic susceptibility value above which
the self-demagnetization must be taken into consideration.
We propose an alternative way of determining this value
based on the body shape and the maximum relative error
allowed in the resultant magnetization.
Our approach is an alternative to the constant value which
seems to be determined empirically and
has been used by the geoscientific community.
Our alternative approach is validated
by the results obtained with numerical simulations.
In a future work,
it would be interesting to use a similar approach to
to bound the maximum relative error in the
magnetic field calculated by neglecting the self-demagnetization.

This work also provide a set of routines to model the magnetic 
field produced by ellipsoids.
The routines are written in Python language as part of
the Fatiando a Terra \citep{uieda2013}, which is an 
open-source library for modelling and inversion in geophysics.
The current version of our code is freely distributed
in a repository of GitHub. We are working to integrate our
routines in the following stable release of Fatiando a Terra.
We hope that these routines be useful to the wide geoscientific
community either in researching and teaching.

@@

The current code is freely distributed, under 
BSD 3-clause licence, in a repository of GitHub
(\url{https://github.com/pinga-lab/magnetic-elipsoid}).
The code was developed as part of the Fatiando a Terra 
\citep{uieda2013}, which is an open-source Python 
library for modelling and inversion in geophysics.
Instructions for running the current version of our code
in the Fatiando a Terra are provided at its repository
in GitHub.
Documentation and installation instructions for 
the most current release version of Fatiando a Terra 
are provided at \url http://www.fatiando.org.
